---
author: "Jacinto Arias"
title: "Get Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Basic usage, report generation

The basic API encapsulates all the functionality of the package in a __single function__, `archeospec::genReport`. It allows different configuration options to generate the combination of outputs carried out in the study.

## Data loading

Data must be provided with an __absolute__ path to directory or tree of directories containing `.asd` files. If you need to configure the data format you can do so by appliying the same parametrization from the `load_signatures_files` files.

For example:

```{r, eval=FALSE}
library(archeospec)

# In windows 10
path <- "C:/Users/(username)/Documents/data/signatures"
# Where signatures contain .asd files and folders with more .asd files


# In linux
path <- "/home/(username)/data/signatures"
```


### Sample data included with the package

We provide sample `.asd` files with the package, you can find it in the package's root folder:

```{r, eval=FALSE}
library(archeospec)

path <- paste(path.package("archeospec"), "/extdata", sep="")   
```


## Experimental approaches

You may choose between different approaches for __unmixing__ and __clustering__ as explored in the original study:

- Unmixing via VCA algorithm
- Unmixing specifiying the endmembers

- Kmeans clustering
- Fixed centroid clustering (selects the endmembers as centroids)

__Unmixing__ can be controlled via the `endmembers` argument, a numeric value will trigger the VCA algorithm and a list of characters values will trigger the manual unmixing, notice that the files must be existing signatures in the data.

__Clustering__ can be controlled via the `kmeans` boolean argument, a positive value will perform the algorithm while a negative one will fix the endmembers.


### Example of VCA + kmeans

```{r archeospec_vca, eval=FALSE}
# Set a seed value to ensure reproducibility
set.seed(934587)

# A default report using VCA to compute k=3 endmembers and kmeans:
genReport(
  input_source=path,
  output="./ex1",
  endmembers=5,
  kmeans=T
)
```

![See rendered result in PDF](report_ex1.pdf){width=0px, height=0px}
[See rendered result in PDF](report_ex1.pdf)

![See rendered result in HTML](report_ex1.html){width=0px, height=0px}
[See rendered result in HTML](report_ex1.html)

### Example of fixed endmembers and clusters

If manual endmembers are selected you may change the displayed name and color for a more clear identification and comparison. This is done via the `endmember_names` and `endmember_colors` property.

```{r archeospec_fixed, eval=FALSE}
# Set a seed value to ensure reproducibility
set.seed(934587)

# A default report usingfixed endmembers and clusters:
genReport(
  input_source=path,
  output="./ex2",
  endmembers=c("signature0000.asd", "signature0001.asd", "signature0002.asd", "signature0003.asd", "signature0005.asd"),
  endmember_names=c("orange", "red", "black", "white", "base"),
  endmember_colors = c("darkorange2", "red2",  "gray15", "white", "green4"),
  kmeans=F
)
```

![See rendered result in PDF](report_ex2.pdf){width=0px, height=0px}
[See rendered result in PDF](report_ex2.pdf)

![See rendered result in HTML](report_ex2.html){width=0px, height=0px}
[See rendered result in HTML](report_ex2.html)


## Data cleaning

We can apply a cleaning process to the data using different options. Please notice that they are __disabled by default__.

- `clean_head` cuts the lower wavelength values until the value specified
- `clean_tail` cuts the higher wavelength values from the value specified
- `clean_leaps` smooth the leaps produced when using multiple sensors for different range of wavelengths, produced between sensor changes. You may pass an arbritary number of points.

```{r archeospec_filter, eval=FALSE}
# Set a seed value to ensure reproducibility
set.seed(934587)

# Filter wavelengths less than 400 and above 2400, and
# smooth the leaps between wavelengths 1000 and 1001, and between 1830 and 1831.
genReport(
  input_source=path,
  output="./ex3",
  endmembers=5,
  kmeans=F,
  clean_head=400,
  clean_tail=2400,
  clean_leaps=c(1001, 1831)
)
```

![See rendered result in PDF](report_ex3.pdf){width=0px, height=0px}
[See rendered result in PDF](report_ex3.pdf)

![See rendered result in HTML](report_ex3.html){width=0px, height=0px}
[See rendered result in HTML](report_ex3.html)


## Controlling the outputs

The report will always include the following plots and tables:

- A brief description of the dataset (list of signatures)
- The list of parameters used to create the report
- The signature plot
- The selected or computed endmembers, in tabular and plot formats
- The elbow plot (if kmeans was performed)
- The correspondence between endmembers and clusters plot
- The endmember weights by cluster bar and whisker box plot
- The residual values for each signature plot
- The summary table for the signatures and its weights by endmembers

Additional plots are included by default, and can be hidden show using the following options:

- `intracorrelation=TRUE` shows the intracorrelation plot. We advise to hide it if not needed as it costly to generate.
- `mutualinfo=TRUE` shows the mutual information between clusters and wavelengths


### Example of hidden plots and named/colored endmembers

```{r archeospec_custom, eval=FALSE}
# Set a seed value to ensure reproducibility
set.seed(934587)

# We might want to customize the report by removing plots which are not interesting for some cases.
genReport(
  input_source=path,
  output="./ex4",
  endmembers=c("signature0000.asd", "signature0001.asd", "signature0002.asd", "signature0003.asd", "signature0005.asd"),
  endmember_names=c("orange", "red", "black", "white", "base"),
  endmember_colors = c("darkorange2", "red2",  "gray15", "white", "green4"),
  kmeans=F,
  clean_head=400,
  clean_tail=2400,
  clean_leaps=c(1001, 1831),
  intracorrelation=F,
  mutualinfo=F
)
```

![See rendered result in PDF](report_ex4.pdf){width=0px, height=0px}
[See rendered result in PDF](report_ex4.pdf)

![See rendered result in HTML](report_ex4.html){width=0px, height=0px}
[See rendered result in HTML](report_ex4.html)


## Controlling the output

`genReport` will produce all different outputs formats by default, if only a particular type of document is required then it can be specified with the `format` options, it takes values from:

- `html_document` (and md)
- `pdf_document` (and LaTeX)
- `word_document`
- `all` (generating all previous document)

```{r archeospec_format, eval=FALSE}
# Set a seed value to ensure reproducibility
set.seed(934587)

# We can render a single format for extended efficiency
genReport(
  input_source=path,
  output="./ex5",
  format="pdf_document"
  endmembers=5,
  kmeans=T
)
```


## API Docs

Please refer to the [documentation] for a comprehensive description of the `genReport` function



