---
author: "Jacinto Arias"
title: "Introduction to archeospec"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Basic usage

#### Create a report composed of elements specified in the parametrization.

First load the data from input source. This must be an ABSOLUTE path to a folder which contains .asd files, or a tree of folders which contains in the end .asd files. These files must contain the data as a pair wavelength<tab>value, and the data must starts after the line file_header (by default "Wavelength<tab>%s").

Then, a filtering process can be applied to the signatures if any of clean_head, clean_tail or clean_leaps are specified. This consists, respectively, on cutting the lower / higher wavelength values. clean_leaps smooth the leaps produced when using multiple sensors for different range of wavelengths, produced between sensor changes.

Finally, the signatures are unmixed. If endmembers is a number, it computes such number of endmembers using the VCA algorithm. If endmembers is a vector of filenames, then those signatures will be selected as endmembers. In this case, we can assign them a representative name and a fixed color (using endmember_names and endmember_colors), so we can easily recognize this signatures in the report. endmember_names is a vector of string names, and endmember_colors is a vector of string colors (check \code{\link{colors}}).


Different outputs can optionaly be included in the report:

-   kmeans: included (TRUE) or not (FALSE). If endmembers is a number, the centroids are chosen randomly. Otherwise, they will be the selected endmembers.
-   intracorrelation: included (TRUE) or not (FALSE).
-   mutualinfo: included (TRUE) or not (FALSE).

The report will always include:

- A brief description of the dataset (list of signatures)
- The list of parameters used to create the report
- The signature plot
- The selected or computed endmembers, in tabular and plot formats
- The elbow plot
- The correspondence between endmembers and clusters plot
- The endmember weights by cluster bar and whisker box plot
- The residual values for each signature plot
- The summary table for the signatures and its weights by endmembers

For more information refer to the documentation and Get starter section in [GitHub](https://github.com/jacintoArias/archeospec/).


```{r archeospec_nofilter, eval=FALSE}
# If we want a report with all the information, using VCA to compute k=3 endmembers:
genReport(
  input_source="/path/to/asd_folder/",
  output="/tmp/archeospec/",
  endmembers=3
  )
```

![Generated report in pdf format](/private/tmp/archeospec/report_files/figure-latex/unnamed-chunk-2-1.pdf)

```{r archeospec_filter, eval=FALSE}
# The same than before, but filtering the signatures: filter out wavelengths less than 400 and above 2400, and
# smooth the leaps between wavelengths 1000 and 1001, and between 1830 and 1831.
genReport(
  input_source="/path/to/asd_folder/",
  output="/tmp/archeospec/",
  endmembers=3,
  clean_head=400,
  clean_tail=2400,
  smooth_points=c(1001, 1831)
  )
```


![Generated report in pdf format](/private/tmp/archeospec/report_files/figure-latex/unnamed-chunk-2-1.pdf)

```{r archeospec_fixed_endmembers, eval=FALSE}
# If we know which signatures are the endmembers, we can fix them:
genReport(
  input_source="/path/to/asd_folder/",
  output="/tmp/archeospec/",
  endmembers=c("almagre.asd.txt", "blanco.asd.txt"),
  clean_head=400,
  clean_tail=2400,
  smooth_points=c(1001, 1831)
  )
```


![Generated report in pdf format](/private/tmp/archeospec/report_files/figure-latex/unnamed-chunk-2-1.pdf)

```{r archeospec_custom, eval=FALSE}
# We might want to customize the report by removing plots which are not interesting for some cases.
genReport(
  input_source="/path/to/asd_folder/",
  output="/tmp/archeospec/",
  endmembers=c("almagre.asd.txt", "blanco.asd.txt"),
  clean_head=400,
  clean_tail=2400,
  smooth_points=c(1001, 1831),
  kmeans=F,
  intracorrelation=F,
  mutualinfo=F
  )
```

![Generated report in pdf format](/private/tmp/archeospec/report_files/figure-latex/unnamed-chunk-2-1.pdf)

