---
title: "Fixed endmembers"
output: html_document
---

```{r setup, include=FALSE}
library(archeospec)
library(purrr)
```

## Loading the data

The package includes a small set of signatures called `signatures`. Load the data

```{r data}
# Loading data
data(signatures)
```


# Fix endmembers

Fix the endmembers to a set of samples. Apply the clustering to group all the signatures to the selected endmembers.

```{r vca_kmeans}
signatures <- unmixing_fixed(
  signatures, 
  files = c("almagre.asd.txt", "negro.asd.txt"),
  names = c("red","black"),
  colors = c("red2", "gray15")
)
signatures <- clustering_endmembers(signatures)
```


# VCA results in tabular format

See information of VCA in tabular format

```{r tables}
table_endmembers(signatures)
table_residuals_summary(signatures)
table_weights(signatures)
```

# VCA and clustering results as plots

See information of VCA and K-Means as plots

```{r plots}
plot_endmembers(signatures)
plot_clusters(signatures)
plot_mutualinfo(signatures)
plot_endmember_cluster(signatures)
plot_endmember_density_bar(signatures)
plot_endmember_density_box(signatures)
plot_residuals(signatures)
```
