---
title: "Fixed endmembers"
output: html_document
---

```{r setup, include=FALSE}
library(archeospec)

# Set a seed value to ensure reproducibility
set.seed(934587)
```

## Loading the data

The package includes a small set of signatures called `signatures`. Load the data

```{r data}
# Loading data
data(signatures)
```


# Fix endmembers

Fix the endmembers to a set of samples. Apply the clustering to group all the signatures to the selected endmembers.

```{r vca_kmeans}
signatures <- unmixing_fixed(
  signatures, 
  endmember_files=c("signature0000.asd", "signature0001.asd", "signature0002.asd", "signature0003.asd", "signature0005.asd"),
  endmember_names=c("orange", "red", "black", "white", "base"),
  endmember_colors = c("darkorange2", "red2",  "gray15", "white", "green4")
)
signatures <- clustering_endmembers(signatures)
```


# VCA results in tabular format

See information of VCA in tabular format

```{r tables}
table_endmembers(signatures)
table_residuals_summary(signatures)
table_weights(signatures)
```

# VCA and clustering results as plots

See information of VCA and K-Means as plots

```{r plots}
plot_endmembers(signatures)
plot_clusters(signatures)
plot_mutualinfo(signatures)
plot_endmember_cluster(signatures)
plot_endmember_density_bar(signatures)
plot_endmember_density_box(signatures)
plot_residuals(signatures)
```
